---
title: "Micro336 Lab 2"
author: "Your Name Here"
date: "7/6/2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

This week's questions and your answers:

# Answer this question before running any R code - read the README in ~/Micro336-master/data/

1. Describe the data you are analyzing today. What was sequenced to generate these samples, and how are these four samples different from one another? (Use your own words to summarize - don't copy and paste the README)

# Answer these two questions using your RQC output
2. 	a. Most Illumina sequencing runs for 16S rRNA are either 150 bp x 2, 250 bp x 2, or 300 bp x 2 paired ends. What is the length of paired end reads used for this Illumina run?

  b. Use the cycle-specific average quality plot to explain whether or not sequencing quality is equally good over the entire read length. Do all samples have equal quality?

  c. What is different about the quality of read1 and read2 in the raw data?
    
  d.  Given that higher read quality indicates a lower sequencing error rate, do read1 or read2 in the raw data have the same error rate?

# Answer this question after you have finished running dada2 and are about to import your data into phyloseq
3. 
a. Did the processed or the raw data lose a greater number of sequences during bioinformatic processing?

b. At which step were the most sequences lost? What happened?

c. How many chimeric sequences were removed?

#Answer part A after you have run chunk 11, part B after you have run chunk 12, and part C using chunk 13.
4
a. How many unique sequences should we find? What method found the most? What method found the fewest? What about for genera?

b. Based on the barplot, can you tell whether the methods agree on which genera are present and how abundant they are?

c. Compare the datasets “RE-Qiime” and “RE_Truth”. How many genera do they both agree on? How many (and which) genera do they disagree about?

# Answer this question after you have used phyloseq to explore your data and compare it to your known truth. Justify your answers.
5. 
a. 	What should the measured abundance of every genus be in the RE community? Does any method achieve this?  Is any method notably better or worse than any other?

b.	Interpret your RS graph. Is any method accurately measuring the abundance of genera? Is any method clearly superior or inferior?

c. Based on your results, is 16S rRNA data an exact representation of actual DNA proportions in a sample?

d. Considering all of your results today (total sequences and genera detected; extra / missing genera, properly measured abundance of taxa), which bioinformatic method performed best overall, and why do you think so? 



```{r, chunk1_setup, echo=FALSE, message=FALSE, warning=FALSE}
#This code chunk loads the libraries needed for today & sets our working directory
library(dada2)
library(Rqc)
library(phyloseq)
library(ggplot2)
library(dplyr)

# Set working directory
setwd("~/Micro336-master/")
#Set path to data files
path <- "~/Micro336-master/data" 
```

Use your lab manual to insert appropriate code for this lab into the corresponding chunks. Answer question #2 after running this chunk.
```{r chunk2_RQC, echo=FALSE, message=FALSE}
# Run RQC to look at quality of reads - insert code here 

```
Insert code to parse forward/reverse reads and parse file names here:
```{r chunk3_fileparse, echo=TRUE, message=FALSE, warnings=FALSE}
```

Insert code to quality-control data prior to running dada2 here:
```{r chunk4_dadaQC, echo=FALSE, message=FALSE}

```
Insert code to run dada2 to here:
```{r chunk5_runDADA, echo=FALSE, message=FALSE}

```
Insert code to make a table and remove chimeras here. 
```{r chunk6_chimeras, echo=FALSE, message=FALSE}

```
Insert code to tally how many sequences made it through your pipeline. Answer question #3 after running this chunk.
```{r chunk7_tally, echo=FALSE, message=FALSE}

```

Insert code to assign taxonomy here
```{r chunk8_taxonomy, echo=FALSE, message=FALSE}

```
Insert your code to create a phyloseq object and summary barplots here. Remember to make barplots at a taxonomic level OTHER THAN PHYLUM here. 
```{r chunk9_physeq, echo=FALSE, message=FALSE}

```

Insert your code to create evaluation object here. 
```{r chunk10_evaluate, echo=FALSE, message=FALSE}

```

Insert your code to evaluate total numbers of sequences (and genera) found by each method. Use these results to answer Question 4a.
```{r chunk11_richness, echo=FALSE, message=FALSE}
```

Insert your code to make barplots summarizing each method here.  Answer question #4b after completing this chunk.
```{r chunk12_barplots, echo=FALSE, message=FALSE}

```

Insert code to calculate unions and intersections between samples. Modify as needed to answer question 4c. Answer question #4c after completing this chunk.
```{r chunk13_intersections, echo=FALSE, message=FALSE}

```

Insert code to compare abundances of individual genera (each dot in the first plot, each column in the second plot) to their measured true abundances (indicated by black dashed lines). Answer question #5 after completing this chunk.
```{r chunk14_finale, echo=FALSE, message=FALSE}

```
