---
title: "Micro336 Lab 2"
author: "Your Name Here"
date: "7/6/2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

This week's questions and your answers:

# Answer this question before running any R code - read the README in ~/Micro336-master/data/

1. Describe the data you are analyzing today. What was sequenced to generate these samples, and how are these four samples different from one another?

# Answer these two questions using your RQC output
2. 	a. Most Illumina sequencing runs for 16S rRNA are either 150 bp x 2, 250 bp x 2, or 300 bp x 2 paired ends. What is the length of paired end reads used for this Illumina run?

  b. Use the cycle-specific average quality plot to explain whether or not sequencing quality is equally good over the entire read length. Do all samples have equal quality?

  c. What is different about the quality of read1 and read2 in the raw data?
    
  e.  Given that higher read quality indicates a lower sequencing error rate, does read1 or read2 in the raw data have a higher error rate?


3.	Use the "Per Read Mean Quality Distribution of Files" and "Read length distributions" plots to answer the following questions:
  a. Which fastq files had the worst quality?

  b. What did the processing step (in which "Raw even" became "processed even" and "Raw staggered" became "Processed staggered") do to the data?

# Answer this question after you have finished running dada2 and are about to import your data into phyloseq
4. 
a. Did the processed or the raw data lose a greater number of sequences during bioinformatic processing?

b. At which step were the most sequences lost?

c. How many chimeric sequences were removed?

# Answer this question after you have used phyloseq to explore your data and compare it to your known truth. Justify your answers.
5. 
a. Did the raw and processed data agree with one another regarding the presence and abundance of taxa within the samples?

b. Based on your answer to A, does how you quality control your data influence your result?

c. How do your staggered and even data compare with the known truth communities? How many total genera were in the original sample, and how many were detected in your data? Were taxa added or missing?

d. Based on your non-log scatterplot of known truth (staggered) vs your data, how well does the abundance of samples in your data reflect its abundance in the original sample? Can you tell whether the processed or raw data comes closer to the truth?

e. Based on your log-axis plots, how well does the abundance of samples in your samples reflect their original abundance? Did the processed or raw data come closer?

f. Based on your results, is 16S rRNA data an exact representation of actual DNA proportions in a sample? 


```{r, chunk1_setup, echo=FALSE, message=FALSE, warning=FALSE}
#This code chunk loads the libraries needed for today & sets our working directory
library("dada2")
library("Rqc")
library("phyloseq")
library("ggplot2")
# Set working directory
setwd("~/Micro336-master/")
#Set path to data files
path <- "~/Micro336-master/data" 
```

Use your lab manual to insert appropriate code for this lab into the corresponding chunks. Answer question #2 after running this chunk.
```{r chunk2_RQC, echo=FALSE, message=FALSE}
# Run RQC to look at quality of reads - insert code here 

```
Insert code to parse forward/reverse reads and parse file names here:
```{r chunk3_fileparse, echo=TRUE, message=FALSE, warnings=FALSE}
```

Insert code to make QC plots with readable X axes here. Answer question #3 after running this chunk.
```{r chunk4_betterplot, echo=FALSE, message=FALSE}

```

Insert code to quality-control data prior to running dada2 here:
```{r chunk5_dadaQC, echo=FALSE, message=FALSE}

```
Insert code to run dada2 to here:
```{r chunk6_runDADA, echo=FALSE, message=FALSE}

```
Insert code to make a table and remove chimeras here. 
```{r chunk7_chimeras, echo=FALSE, message=FALSE}

```
Insert code to tally how many sequences made it through your pipeline. Answer question #4 after running this chunk.
```{r chunk8_tally, echo=FALSE, message=FALSE}

```

Insert code to assign taxonomy here
```{r chunk9_taxonomy, echo=FALSE, message=FALSE}

```
Insert your code to create a phyloseq object and summary barplots here. Remember to make barplots at a taxonomic level OTHER THAN PHYLUM here. 
```{r chunk10_physeq, echo=FALSE, message=FALSE}

```

Insert your code to create evaluation object here. 
```{r chunk11_evaluate, echo=FALSE, message=FALSE}

```

Insert your code to evaluate unions and intersections here
```{r chunk12_sets, echo=FALSE, message=FALSE}
```

Insert your code to plot results vs known truths here. Answer question #5 after completing this chunk.
```{r chunk13_evalplots, echo=FALSE, message=FALSE}
```

