---
title: "Micro336 Lab 4"
author: "SEM"
date: "September 12, 2017"
output:
  html_document: default
pdf_document: default
---
  
  This week's questions and your answers:
  
1. Create a new dataset from the HMP data, containing only Throat samples sequenced at JCVI. Remember to rarefy (1000 sequences per sample) and trim (absent taxa removed) your dataset from last week. Show a plot with changes in Phyla by sex using only groups representing more than 10% of data of relative abundance.

#Your code to load, subset, and trim data here.
```{r}

``` 

# Your code to plot abundances here
```{r}

```

2. Identify one phylum which may be affected by sex. Subset your dataset and only plot that phylum.

# Your code to subset data by a particular phylum, prune out genera with no hits, and plot 
```{r}

```

3. Re-plot data using absolute abundance and facet plots by Genera

#Your code of data faceted by genera here.
```{r}

```


4. 
a. What is the average dissimilarity between males and females in your dataset?
b. How many OTUs make up the top 15% of the variance in the generated SIMPER file?

#Your code to do a SIMPER analysis on difference between males and females here
```{r}

``` 


5. Provide a plot showing the Phyla represented by organisms identified in SIMPER analysis contributing the top 30% of the variance.

#Your code for this plot here:
```{r}


``` 

6. Provide a plot showing the genera represented by organisms identified in SIMPER analysis contributing the top 30% of the variance. 

#Your code here
```{r}

```


7. Provide a facetted plot by genus with error bars for Genera identified in SIMPER analysis.

#Your code here
```{r, fig.width=12}

```

8. 
a. Which genera seem to show distinct abundances due to sex?

b. How would you confirm if this was true?



