---
  title: "Micro336 Week 3"
author: "SEM"
date: "August 12, 2021"
output:
  html_document: default
pdf_document: default
---
  
  NOTE: Remember to rename your file so that it follows this format -> StudentID_Lastname_Week3
  
  This week's questions and your answers:
  
1. Provide a Stacked barplot showing mean relative abundance of phyla in Throat samples based on RUNCENTER
```{r Q1}

``` 


2. Provide a plot showing differentialy abundant taxa between RUNCENTER by Class and Phylum for Throat samples. Indicate which Taxa are different.

```{r Q2}

``` 

3. Discuss which plot (Q1 or 2) is better and why.


4. What other approach could you use to identify taxa changing by RUNCENTER? 


5. Plot differentialy abundant taxa by Genus and Phylum as determined using approach in Q4.

```{r}


```









If your results won't knit use the commands below to create your final document. 
##library(knitr)
##knit('Week3.rmd')
###will make an output file called file.rm
##library(rmarkdown)
##render("Week3.rmd","html_document")
